Early Warning Signals and the Prosecutor's Fallacy

### 2012-09-06 prosecutor/fallacy.tex
(re)submitted copies to PRSB

%\documentclass[authoryear,5p]{elsarticle}
\documentclass[authoryear,review,12pt]{elsarticle}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
\usepackage{amsmath,amsfonts}
\usepackage{lineno}
\linenumbers
%\usepackage{float}
\usepackage{subfigure}
%\usepackage[pdftex]{color}
%\definecolor{darkblue}{rgb}{0,0,0.5}
%\definecolor{darkgreen}{rgb}{0,0.5,0}
%\usepackage[pdftex, colorlinks]{hyperref}
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\newcommand{\ud}{\mathrm{d}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\C}{\mathrm{Cov}}
\newcommand{\V}{\mathrm{Var}}
\newcommand{\cb}[1]{{\it \color{darkgreen} (#1)}}

@@

\usepackage{endfloat} % Move floats to the end of the document
\renewcommand{\listoffigures}{} % but supress these lists
\renewcommand{\listoftables}{} % supress these lists
% compile without figures?
\newif\ifhavefigures
\havefigurestrue
\newif\ifnohavefigures % allows alternate condition
%\nohavefigurestrue     % turn on the alternate condition uncommenting this

@@

\begin{figure}
  \begin{center}
      \ifhavefigures
        \includegraphics[width=\textwidth]{Figure1}
      \fi
      \ifnohavefigures
        \hspace{\textwidth}
      \fi
      \label{fig:bifurcation}
  \end{center} \caption{\textbf{The Prosecutor's Fallacy}. 
\textbf{(a)} Plot of the model functions shown in Eq~\eqref{master}
with parameters $a=180$, $K=500$, $e=.5$, and $h=200$.  When the death
rate is higher than the birth rate, the system dynamics drive the state
(population size) to smaller values.  When birth rate is higher, the
system moves right, as indicated by the arrows.\textbf{(b)} The potential
energy is given by the negative integral of $b(n)-d(n)$, shown in the
lower plot.   The potential function gives an intuitive picture of
the stability of a system by imagining the curve as a surface on which
a ball is free to bounce across -- wells correspond to stable points
and peaks to unstable points. While most trajectories remain near the
stable well, some transition out merely by chance.  An example of such
a trajectory is shown in the top panel, in which time increases along
the vertical axis. Though initially oscillating around the stable state,
a chance excursion carries it beyond the threshold (vertical dotted line).
Such chance trajectories can produce the statistical patterns as observed
in true critical transitions seen in panel \textbf{(c)}: Early warning
signals  are aimed at detecting systems which are slowly moving towards
a tipping point or bifurcation, illustrated in the successive curves
(detiorating and critical). Top panel: An example trajectory from a
simulation under this process shows the state of the system as the
potential moves towards the bifurcation point.
} \label{fig:1}
\end{figure}

@@

To demonstrate the effect we observe is not unique to models with Allee
effects, we provide an example of the effect arising in a discrete-time
model with two non-zero stable states adapted from~\citep{May1977},
\begin{equation}
X_{t+1} = X_t \exp\left( r \left(1 - \frac{ X_t }{ K } \right) - \frac{ a * X_t ^ {Q - 1} }{s ^ Q + H ^ Q} \right). \label{May1977}
\end{equation}
Annotated, parallel code to replicate the analysis presented here is
also provided in the appendix.

\begin{figure}
  \begin{center}
    \ifhavefigures
    \includegraphics[width=6in]{figure2}
    \fi
  \end{center}
  \caption{The distribution of the correlation statistic $\tau$ for two
  early warning indicators (variance, autocorrelation) on replicates
  conditionally selected for having collapsed by chance in simulations
  is shown in grey bars.  Solid lines indicate the estimated density of
  the statistic from a random sample of the simulations (not conditional
  on observing a transition). Positive values of $\tau$ correspond to
  a pattern of an indicator increasing with time; typically taken as
  evidence that a system is approaching a critical transition.  In these
  simulations, the pattern arises instead from the Prosecutor's fallacy
  of conditional selection.}
  \label{fig:indicator}
\end{figure}

@@

\section{References}%bibliography
\bibliography{fallacy}

 \end{document}

 \section{References}%bibliography

 \begin{thebibliography}{24}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

\bibitem[{Berkes et~al.(2006)Berkes, Hughes, Steneck, Wilson, Bellwood, Crona,
  Folke, Gunderson, Leslie, Norberg, Nystr\"{o}m, Olsson, Osterblom, Scheffer,
  and Worm}]{Berkes2006}
Berkes, F., Hughes, T.~P., Steneck, R.~S., Wilson, J.~A., Bellwood, D.~R.,
  Crona, B., Folke, C., Gunderson, L.~H., Leslie, H.~M., Norberg, J.,
  Nystr\"{o}m, M., Olsson, P., Osterblom, H., Scheffer, M., Worm, B., Mar.
  2006. {Ecology. Globalization, roving bandits, and marine resources.} Science
  (New York, N.Y.) 311~(5767), 1557--8.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/16543444}

\bibitem[{Boettiger and Hastings(2012)}]{Boettiger2012b}
Boettiger, C., Hastings, A., 2012. {Quantifying Limits to Detection of Early
  Warning for Critical Transitions}. Journal Of The Royal Society Interface.

\bibitem[{{California Supreme Court}(1968)}]{PeopleCollins1968}
{California Supreme Court}, 1968. People v. collins.

\bibitem[{Carpenter(2011)}]{Carpenter2011}
Carpenter, J., Feb. 2011. {May the Best Analyst Win}. Science 331~(6018),
  698--699.
\newline\urlprefix\url{http://www.sciencemag.org/cgi/doi/10.1126/science.331.6018.698}

\bibitem[{Carpenter and Brock(2006)}]{Carpenter2006}
Carpenter, S.~R., Brock, W.~A., 2006. {Rising variance: a leading indicator of
  ecological transition.} Ecology letters 9~(3), 311--8.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/16958897}

\bibitem[{Courchamp et~al.(2008)Courchamp, Berec, and
  Gascoigne}]{Courchamp2008}
Courchamp, F., Berec, L., Gascoigne, J., 2008. {Allee Effects in Ecology and
  Conservation}. Oxford University Press, USA.
\newline\urlprefix\url{http://www.amazon.com/Effects-Ecology-Conservation-Franck-Courchamp/dp/0198570309}

\bibitem[{Dai et~al.(2012)Dai, Vorselen, Korolev, and Gore}]{Dai2012}
Dai, L., Vorselen, D., Korolev, K.~S., Gore, J., May 2012. {Generic Indicators
  for Loss of Resilience Before a Tipping Point Leading to Population
  Collapse}. Science 336~(6085), 1175--1177.
\newline\urlprefix\url{http://www.sciencemag.org/cgi/doi/10.1126/science.1219805}

@@

\bibitem[{Dakos et~al.(2008)Dakos, Scheffer, van Nes, Brovkin, Petoukhov, and
  Held}]{Dakos2008}
Dakos, V., Scheffer, M., van Nes, E.~H., Brovkin, V., Petoukhov, V., Held, H.,
  Sep. 2008. {Slowing down as an early warning signal for abrupt climate
  change.} Proceedings of the National Academy of Sciences 105~(38), 14308--12.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/18787119}

\bibitem[{Ditlevsen and Johnsen(2010)}]{Ditlevsen2010}
Ditlevsen, P.~D., Johnsen, S.~J., Oct. 2010. {Tipping points: Early warning and
  wishful thinking}. Geophysical Research Letters 37~(19), 2--5.
\newline\urlprefix\url{http://www.agu.org/pubs/crossref/2010/2010GL044486.shtml}

\bibitem[{Drake and Griffen(2010)}]{Drake2010}
Drake, J.~M., Griffen, B.~D., Sep. 2010. {Early warning signals of extinction
  in deteriorating environments}. Nature 467~(7314), 456--459.
\newline\urlprefix\url{http://www.nature.com/doifinder/10.1038/nature09389}

\bibitem[{Guttal and Jayaprakash(2008)}]{Guttal2008}
Guttal, V., Jayaprakash, C., Dec. 2008. {Spatial variance and spatial skewness:
  leading indicators of regime shifts in spatial ecological systems}.
  Theoretical Ecology 2~(1), 3--12.
\newline\urlprefix\url{http://www.springerlink.com/index/10.1007/s12080-008-0033-1}

\bibitem[{Hirota et~al.(2011)Hirota, Holmgren, {Van Nes}, and
  Scheffer}]{Hirota2011}
Hirota, M., Holmgren, M., {Van Nes}, E.~H., Scheffer, M., Oct. 2011. {Global
  Resilience of Tropical Forest and Savanna to Critical Transitions}. Science
  334~(6053), 232--235.
\newline\urlprefix\url{http://www.sciencemag.org/cgi/doi/10.1126/science.1210657}

\bibitem[{K\'{e}fi et~al.(2007)K\'{e}fi, Rietkerk, Alados, Pueyo, Papanastasis,
  Elaich, and de~Ruiter}]{Kefi2007}
K\'{e}fi, S., Rietkerk, M., Alados, C.~L., Pueyo, Y., Papanastasis, V.~P.,
  Elaich, A., de~Ruiter, P.~C., Sep. 2007. {Spatial vegetation patterns and
  imminent desertification in Mediterranean arid ecosystems.} Nature
  449~(7159), 213--7.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/17851524}

\bibitem[{Lenton(2011)}]{Lenton2011}
Lenton, T.~M., Jun. 2011. {Early warning of climate tipping points}. Nature
  Climate Change 1~(4), 201--209.
\newline\urlprefix\url{http://www.nature.com/doifinder/10.1038/nclimate1143}

\bibitem[{Lenton et~al.(2012)Lenton, Livina, and Dakos}]{Lenton2012}
Lenton, T.~M., Livina, V., Dakos, V., 2012. {Early warning of climate tipping
  points from critical slowing down: comparing methods to improve robustness}.
  Philosophical Transactions of The Royal Society A (in press).
\newline\urlprefix\url{http://sites.google.com/site/noraclpreprints/contributions/lenton-et-al/Early\_warning\_methods\_revised\_with\_figs\_and\_appendix2.pdf}

\bibitem[{Livina and Lenton(2007)}]{Livina2007}
Livina, V.~N., Lenton, T.~M., Feb. 2007. {A modified method for detecting
  incipient bifurcations in a dynamical system}. Geophysical Research Letters
  34~(3), 1--5.
\newline\urlprefix\url{http://www.agu.org/pubs/crossref/2007/2006GL028672.shtml}

\bibitem[{Mumby et~al.(2007)Mumby, Hastings, and Edwards}]{Mumby2007}
Mumby, P.~J., Hastings, A., Edwards, H.~J., Nov. 2007. {Thresholds and the
  resilience of Caribbean coral reefs.} Nature 450~(7166), 98--101.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/17972885}

\bibitem[{Scheffer et~al.(2009)Scheffer, Bascompte, Brock, Brovkin, Carpenter,
  Dakos, Held, van Nes, Rietkerk, and Sugihara}]{Scheffer2009}
Scheffer, M., Bascompte, J., Brock, W.~A., Brovkin, V., Carpenter, S.~R.,
  Dakos, V., Held, H., van Nes, E.~H., Rietkerk, M., Sugihara, G., 2009.
  {Early-warning signals for critical transitions.} Nature 461~(7260), 53--9.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/19727193}

\bibitem[{Scheffer et~al.(2001)Scheffer, Carpenter, Foley, Folke, and
  Walker}]{Scheffer2001}
Scheffer, M., Carpenter, S.~R., Foley, J.~A., Folke, C., Walker, B., Oct. 2001.
  {Catastrophic shifts in ecosystems.} Nature 413~(6856), 591--6.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/11595939}

\bibitem[{Thompson and Sieber(2010)}]{Thompson2010}
Thompson, J. M.~T., Sieber, J., Dec. 2010. {Climate tipping as a noisy
  bifurcation: a predictive technique}. IMA Journal of Applied Mathematics
  76~(1), 27--46.
\newline\urlprefix\url{http://imamat.oxfordjournals.org/cgi/doi/10.1093/imamat/hxq060}

\bibitem[{Thompson and Schumann(1987)}]{Thompson1987}
Thompson, W.~C., Schumann, E.~L., 1987. {Interpretation of statistical evidence
  in criminal trials: The prosecutor's fallacy and the defense attorney's
  fallacy.} Law and Human Behavior 11~(3), 167--187.
\newline\urlprefix\url{http://psycnet.apa.org/journals/lhb/11/3/167/
  http://doi.apa.org/getdoi.cfm?doi=10.1007/BF01044641}

\bibitem[{van Nes and Scheffer(2007)}]{vanNes2007}
van Nes, E.~H., Scheffer, M., Jun. 2007. {Slow recovery from perturbations as a
  generic indicator of a nearby catastrophic shift.} The American naturalist
  169~(6), 738--47.
\newline\urlprefix\url{http://www.ncbi.nlm.nih.gov/pubmed/17479460}

\bibitem[{Veraart et~al.(2011)Veraart, Faassen, Dakos, van Nes, L\"{u}rling,
  and Scheffer}]{Veraart2011}
Veraart, A.~J., Faassen, E.~J., Dakos, V., van Nes, E.~H., L\"{u}rling, M.,
  Scheffer, M., Dec. 2011. {Recovery rates reflect distance to a tipping point
  in a living system}. Nature, 2--5.
\newline\urlprefix\url{http://www.nature.com/doifinder/10.1038/nature10723}

\end{thebibliography}

>>>>

%\documentclass[authoryear,5p]{elsarticle}
\documentclass[authoryear,review,12pt]{elsarticle}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
\usepackage{amsmath,amsfonts}
\usepackage{lineno}
\linenumbers
\usepackage{subfigure}
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\newcommand{\ud}{\mathrm{d}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\C}{\mathrm{Cov}}
\newcommand{\V}{\mathrm{Var}}
\newcommand{\cb}[1]{{\it \color{darkgreen} (#1)}}

@@

\usepackage{endfloat} % Move floats to the end of the document
\renewcommand{\listoffigures}{} % but supress these lists
\renewcommand{\listoftables}{} % supress these lists
% compile without figures?
\newif\ifhavefigures
\havefigurestrue
\newif\ifnohavefigures % allows alternate condition
\nohavefigurestrue     % turn on the alternate condition uncommenting this

@@

\begin{figure}
  \begin{center}
      \ifhavefigures
%        \includegraphics[width=\textwidth]{Figure1}
      \fi
      \ifnohavefigures
        \hspace{\textwidth}
      \fi
      \label{fig:bifurcation}
  \end{center} \caption{\textbf{The Prosecutor's Fallacy}. 
\textbf{(a)} Plot of the model functions shown in Eq~\eqref{master}
with parameters $a=180$, $K=500$, $e=.5$, and $h=200$.  When the death
rate is higher than the birth rate, the system dynamics drive the state
(population size) to smaller values.  When birth rate is higher, the
system moves right, as indicated by the arrows.\textbf{(b)} The potential
energy is given by the negative integral of $b(n)-d(n)$, shown in the
lower plot.   The potential function gives an intuitive picture of
the stability of a system by imagining the curve as a surface on which
a ball is free to bounce across -- wells correspond to stable points
and peaks to unstable points. While most trajectories remain near the
stable well, some transition out merely by chance.  An example of such
a trajectory is shown in the top panel, in which time increases along
the vertical axis. Though initially oscillating around the stable state,
a chance excursion carries it beyond the threshold (vertical dotted line).
Such chance trajectories can produce the statistical patterns as observed
in true critical transitions seen in panel \textbf{(c)}: Early warning
signals  are aimed at detecting systems which are slowly moving towards
a tipping point or bifurcation, illustrated in the successive curves
(detiorating and critical). Top panel: An example trajectory from a
simulation under this process shows the state of the system as the
potential moves towards the bifurcation point.
} \label{fig:1}
\end{figure}

@@

To demonstrate the effect we observe is not unique to models with Allee
effects, we provide an example of the effect arising in a discrete-time
model with two non-zero stable states adapted from~\citep{May1977},
\begin{equation}
X_{t+1} = X_t \exp\left( r \left(1 - \frac{ X_t }{ K } \right) - \frac{ a * X_t ^ {Q - 1} }{X_t ^ Q + H ^ Q} \right). \label{May1977}
\end{equation}
which combines a logistic growth model with a saturating predator response (See~\citet{May1977} for detailed discussion), shown in Figure~\ref{fig:may}.  Code to replicate the analysis can be found at \\ 
\verb|https://github.com/cboettig/earlywarning/tree/prosecutor/|.

\begin{figure}
  \begin{center}
    \ifhavefigures
%    \includegraphics[width=6in]{Figure2}
    \fi
     \ifnohavefigures
     \hspace{6in}
    \fi
 \end{center}
  \caption{The distribution of the correlation statistic $\tau$ for two
  early warning indicators (variance, autocorrelation) on replicates
  conditionally selected for having collapsed by chance in simulations
  is shown in grey bars.  Solid lines indicate the estimated density of
  the statistic from a random sample of the simulations (not conditional
  on observing a transition). Positive values of $\tau$ correspond to
  a pattern of an indicator increasing with time; typically taken as
  evidence that a system is approaching a critical transition.  In these
  simulations, the pattern arises instead from the Prosecutor's fallacy
  of conditional selection.}
  \label{fig:indicator}
\end{figure}

\begin{figure}
  \begin{center}
    \ifhavefigures
%    \includegraphics[width=6in]{Figure3}
    \fi
    \ifnohavefigures
     \hspace{6in}
    \fi
  \end{center}
  \caption{The identical analysis from Figure~\ref{fig:indicator} is shown for the model in Eq~\eqref{May1977} using parameters $r = 0.75$, $K=10$, $a=1.7$, $Q=3$, and $H=1$. A similar statistical bias, particularly towards positive values of $tau$ occurs in this model as well.}
  \label{fig:may}
\end{figure}

@@

\begin{thebibliography}{25}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi

\bibitem[{Berkes et~al.(2006)Berkes, Hughes, Steneck, Wilson, Bellwood, Crona,
  Folke, Gunderson, Leslie, Norberg, Nystr\"{o}m, Olsson, Osterblom, Scheffer,
  and Worm}]{Berkes2006}
Berkes, F., Hughes, T.~P., Steneck, R.~S., Wilson, J.~A., Bellwood, D.~R.,
  Crona, B., Folke, C., Gunderson, L.~H., Leslie, H.~M., Norberg, J.,
  Nystr\"{o}m, M., Olsson, P., Osterblom, H., Scheffer, M., Worm, B., Mar.
  2006. {Ecology. Globalization, roving bandits, and marine resources.} Science
  (New York, N.Y.) 311~(5767), 1557--8.

\bibitem[{Boettiger and Hastings(2012)}]{Boettiger2012b}
Boettiger, C., Hastings, A., May 2012. {Quantifying limits to detection of
  early warning for critical transitions}. Journal of The Royal Society
  Interface 9~(75), 2527 -- 2539.

\bibitem[{Carpenter(2011)}]{Carpenter2011}
Carpenter, J., Feb. 2011. {May the Best Analyst Win}. Science 331~(6018),
  698--699.

\bibitem[{Carpenter and Brock(2006)}]{Carpenter2006}
Carpenter, S.~R., Brock, W.~A., 2006. {Rising variance: a leading indicator of
  ecological transition.} Ecology letters 9~(3), 311--8.

\bibitem[{Courchamp et~al.(2008)Courchamp, Berec, and
  Gascoigne}]{Courchamp2008}
Courchamp, F., Berec, L., Gascoigne, J., 2008. {Allee Effects in Ecology and
  Conservation}. Oxford University Press, USA.

\bibitem[{Dai et~al.(2012)Dai, Vorselen, Korolev, and Gore}]{Dai2012}
Dai, L., Vorselen, D., Korolev, K.~S., Gore, J., May 2012. {Generic Indicators
  for Loss of Resilience Before a Tipping Point Leading to Population
  Collapse}. Science 336~(6085), 1175--1177.

@@

\bibitem[{Dakos et~al.(2008)Dakos, Scheffer, van Nes, Brovkin, Petoukhov, and
  Held}]{Dakos2008}
Dakos, V., Scheffer, M., van Nes, E.~H., Brovkin, V., Petoukhov, V., Held, H.,
  Sep. 2008. {Slowing down as an early warning signal for abrupt climate
  change.} Proceedings of the National Academy of Sciences 105~(38), 14308--12.

\bibitem[{Ditlevsen and Johnsen(2010)}]{Ditlevsen2010}
Ditlevsen, P.~D., Johnsen, S.~J., Oct. 2010. {Tipping points: Early warning and
  wishful thinking}. Geophysical Research Letters 37~(19), 2--5.

\bibitem[{Drake and Griffen(2010)}]{Drake2010}
Drake, J.~M., Griffen, B.~D., Sep. 2010. {Early warning signals of extinction
  in deteriorating environments}. Nature 467~(7314), 456--459.

\bibitem[{Guttal and Jayaprakash(2008)}]{Guttal2008}
Guttal, V., Jayaprakash, C., Dec. 2008. {Spatial variance and spatial skewness:
  leading indicators of regime shifts in spatial ecological systems}.
  Theoretical Ecology 2~(1), 3--12.

\bibitem[{Hirota et~al.(2011)Hirota, Holmgren, {Van Nes}, and
  Scheffer}]{Hirota2011}
Hirota, M., Holmgren, M., {Van Nes}, E.~H., Scheffer, M., Oct. 2011. {Global
  Resilience of Tropical Forest and Savanna to Critical Transitions}. Science
  334~(6053), 232--235.

\bibitem[{K\'{e}fi et~al.(2007)K\'{e}fi, Rietkerk, Alados, Pueyo, Papanastasis,
  Elaich, and de~Ruiter}]{Kefi2007}
K\'{e}fi, S., Rietkerk, M., Alados, C.~L., Pueyo, Y., Papanastasis, V.~P.,
  Elaich, A., de~Ruiter, P.~C., Sep. 2007. {Spatial vegetation patterns and
  imminent desertification in Mediterranean arid ecosystems.} Nature
  449~(7159), 213--7.

\bibitem[{Lenton(2011)}]{Lenton2011}
Lenton, T.~M., Jun. 2011. {Early warning of climate tipping points}. Nature
  Climate Change 1~(4), 201--209.

\bibitem[{Lenton et~al.(2012)Lenton, Livina, and Dakos}]{Lenton2012}
Lenton, T.~M., Livina, V., Dakos, V., 2012. {Early warning of climate tipping
  points from critical slowing down: comparing methods to improve robustness}.
  Philosophical Transactions of The Royal Society A (in press).

\bibitem[{Livina and Lenton(2007)}]{Livina2007}
Livina, V.~N., Lenton, T.~M., Feb. 2007. {A modified method for detecting
  incipient bifurcations in a dynamical system}. Geophysical Research Letters
  34~(3), 1--5.

\bibitem[{May(1977)}]{May1977}
May, R.~M., Oct. 1977. {Thresholds and breakpoints in ecosystems with a
  multiplicity of stable states}. Nature 269~(5628), 471--477.

\bibitem[{Mumby et~al.(2007)Mumby, Hastings, and Edwards}]{Mumby2007}
Mumby, P.~J., Hastings, A., Edwards, H.~J., Nov. 2007. {Thresholds and the
  resilience of Caribbean coral reefs.} Nature 450~(7166), 98--101.

\bibitem[{Scheffer et~al.(2009)Scheffer, Bascompte, Brock, Brovkin, Carpenter,
  Dakos, Held, van Nes, Rietkerk, and Sugihara}]{Scheffer2009}
Scheffer, M., Bascompte, J., Brock, W.~A., Brovkin, V., Carpenter, S.~R.,
  Dakos, V., Held, H., van Nes, E.~H., Rietkerk, M., Sugihara, G., 2009.
  {Early-warning signals for critical transitions.} Nature 461~(7260), 53--9.

\bibitem[{Scheffer et~al.(2001)Scheffer, Carpenter, Foley, Folke, and
  Walker}]{Scheffer2001}
Scheffer, M., Carpenter, S.~R., Foley, J.~A., Folke, C., Walker, B., Oct. 2001.
  {Catastrophic shifts in ecosystems.} Nature 413~(6856), 591--6.

\bibitem[{{Supreme Court}(1968)}]{PeopleCollins1968}
{Supreme Court}, C., 1968. {People v. Collins}.

\bibitem[{Thompson and Sieber(2010)}]{Thompson2010}
Thompson, J. M.~T., Sieber, J., Dec. 2010. {Climate tipping as a noisy
  bifurcation: a predictive technique}. IMA Journal of Applied Mathematics
  76~(1), 27--46.

\bibitem[{Thompson and Schumann(1987)}]{Thompson1987}
Thompson, W.~C., Schumann, E.~L., 1987. {Interpretation of statistical evidence
  in criminal trials: The prosecutor's fallacy and the defense attorney's
  fallacy.} Law and Human Behavior 11~(3), 167--187.

\bibitem[{van Nes and Scheffer(2007)}]{vanNes2007}
van Nes, E.~H., Scheffer, M., Jun. 2007. {Slow recovery from perturbations as a
  generic indicator of a nearby catastrophic shift.} The American naturalist
  169~(6), 738--47.

\bibitem[{Veraart et~al.(2011)Veraart, Faassen, Dakos, van Nes, L\"{u}rling,
  and Scheffer}]{Veraart2011}
Veraart, A.~J., Faassen, E.~J., Dakos, V., van Nes, E.~H., L\"{u}rling, M.,
  Scheffer, M., Dec. 2011. {Recovery rates reflect distance to a tipping point
  in a living system}. Nature, 2--5.

\end{thebibliography}
 \end{document}
